---
title: "Gif create rmd"
author: "A Parker + DVM Bishop"
date: "03/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("gganimate")
library("beeswarm")
library("dplyr")
library("animation")
library("FSA")
library("data.table")
library("ggbeeswarm")
library("effsize")
library("BayesFactor")
library(gtools) # for chr
library(permute) #for shuffling effect sizes

source("split_vio.R")
set.seed(5)#make data reproducible
options(scipen = 999) #turn off sci notation
```

## Explanation

For the Gorilla game, we need gifs that show how a dataset changes as additional datapoints are sampled. This scripts creates gifs. We also create the spreadsheet that will be used to control display of trials in the Gorilla game, and we will save the data behind the gifs so we can check how participants' responses relate to the observed effect size, etc.

```{r num2alpha.function}
# convert number to letter pair for values > 26
# after A to Z, we code runs as AA, AB etc (like Excel cols)
# This is only needed when making a v large simulated dataset. 

num2alpha<-function(n){
  i<-as.integer((n-1)/26)
  pre<-''
  if(i>0){
    pre<-chr(64+i)
  }
  post<-chr(64+n%%26)
if(post=='@'){post<-'Z'}
  myletters<-paste0(pre,post)
  return(myletters)
}
```

## Prepare to simulate
Stimuli will be presented in blocks, each of which contains a mixture of two effect sizes - either zero (null) or E (true effect). We will start with E = .3. 


```{r prepsim}
gifmake<-1 #set to 1 for creating gifs for study; 0 if already made
Elist<-c(3,0) #effect size x 10

nsets<-c(24,24) #samples when doing gifs
blocklength<-12 #sum of nsets should be divisible exactly by blocklength
ntrial<-sum(nsets)

N_size <- c(20, 40, 80, 160, 320, 640) #sample sizes to use
allN<-length(N_size)
N <-max(N_size) #on each run simulate max number
thistrial<-0 #initialise counter 

 #make spreadsheet for Gorilla
myspread<-data.frame(matrix(nrow=(4+length(nsets)+sum(nsets)),ncol=60)) #rows need to allow for intro & end(3) breaks (length(nsets))+row for each trial (sum(nsets))
names(myspread)<-c('trial','randomise_blocks','randomise_trials','display','ANSWER','image1','image1f','ES','EarningCorrect','EarningWrong',
                   'skip_ref',
                   'meanC1','sdC1','meanE1','sdE1',
                   'meanC2','sdC2','meanE2','sdE2',
                   'meanC3','sdC3','meanE3','sdE3',
                   'meanC4','sdC4','meanE4','sdE4',
                   'meanC5','sdC5','meanE5','sdE5',
                   'meanC6','sdC6','meanE6','sdE6',
                   'ES1','ES2','ES3','ES4','ES5','ES6',
                   'LL1','LL2','LL3','LL4','LL5','LL6',
                   't1','t2','t3','t4','t5','t6',
                    'BF1','BF2','BF3','BF4','BF5','BF6')

# Prepopulate spreadsheet
myspread$display<-'beeswarms' #default
myspread$display[1]<-'instructions'
myspread$display[2]<-'demo'

myspread$image1[1:2]<-'E5_demo.gif' #predetermined gifs for demo
myspread$image1f[1:2]<-'E5_demo.png'

myspread$EarningCorrect<-10
myspread$EarningWrong<- (-10)

myspread$ANSWER <-'Blue=Pink' #default


#make dataframe to save simulated data for each run

mydf<-data.frame(matrix(nrow=(N),ncol=2))
colnames(mydf)<-c('Group','Score')
mydf$Group <- 'Experimental'
mydf$Group[1:N/2] <- 'Control'
mydf$col<-2 #red and blue dots
mydf$col[1:N/2]<-4

```


```{r makegif}
makegif<-function(big_data,E,filename) {
  
p <- ggplot(big_data, aes(x= cnd_sample, y= Score, color= Group)) +
  geom_beeswarm(cex = 1.5, dodge.width = .2) + 
  xlab("Sample size per condition") + ylim(-3,3)
p <- p + theme_bw() + theme(legend.title = element_blank()) + theme(legend.position="top") +
  theme(text = element_text(size=15))
#p
# set animation values
anim <- p + transition_states(cnd_sample,
              transition_length = 10,
              state_length = 60) +
            enter_grow() + exit_shrink()
anim <- animate(anim, fps=1, duration=12, renderer= gifski_renderer(loop = F)) #12 frames,2 of each N
# print animation
#anim

# save,
filename2<-paste0('gifs_pngs/',filename,'.gif')
anim_save(filename2, anim)
# answer plot ==================================================================================================

truedat <- big_data


# create means for horizontal lines
meanC<-0
meanE<-E

#Just use big_data to provide data frame, but we'll just create new values in a v big sample

truedat<-rbind(big_data,big_data,big_data,big_data)
range1<-1:(nrow(truedat)/2)
range2<-((nrow(truedat)/2)+1):nrow(truedat)
truedat$Group<-'Experimental'
truedat$Group[range1]<-'Control'
truedat$Score[range1]<-rnorm(length(range1),0,1)
truedat$Score[range2]<-rnorm(length(range1),E,1)
truedat$col<-2
truedat$col[range1]<-4
truedat$cnd_sample<-10000 # default is ultralarge sample
truedat$cnd_sample[1:(allN-1)]<-N_size[1:(allN-1)] #retain some points coded with N_size so plot shows this range
truedat$Score[1:allN]<-(-1000) #won't display
truedat$cnd_sample<-as.factor(truedat$cnd_sample)


# create answer plot
q <- ggplot(truedat, aes(x= cnd_sample, y= Score, color= Group, fill= Group)) +
  geom_hline(yintercept = 0, linetype="dotdash", color = "#F8766D", size=1) + 
  geom_hline(yintercept = E, linetype="longdash", color = "#00BFC4", size=1) + 
  geom_split_violin(trim= TRUE, alpha= 0.5, width= -1) +
  geom_boxplot(width = 0.1, position = position_dodge(width=0.25), outlier.alpha = 0) + 
  xlab("Sample size per condition") + ylab("Score") + ylim(-3,3)
q <- q + theme_bw() + theme(legend.title = element_blank()) + theme(legend.position="top")  +
  theme(text = element_text(size=15))
q
filename3<-paste0('gifs_pngs/',filename,'.png')
ggsave(filename3, dpi= 300, width= 7, height = 7)  

return()
}

```

```{r runloop}
# Make a list of effect sizes and shuffle them so tht there is a mixture in each block
tempr<-c(rep(Elist[1],nsets[1]),rep(Elist[2],nsets[2])) #all effect size values as a long list
trialE <-tempr[shuffle(tempr)] #all effect sizes in random order
myblock<-1
blockcount<-0
spreadrowcount <- 2 #keep count of row number for writing to spreadsheet - will start at row 3, after demo and instructions
#start loop here

  for (thistrial in 1:ntrial){

    E <- trialE[thistrial]/10 #select the next effect size from the trialE vector (ie shuffled values)
    spreadrowcount<-spreadrowcount+1
    blockcount<-blockcount+1
    if(blockcount>blocklength)
       {blockcount<-1
         myblock<-myblock+1
         myspread[spreadrowcount,]<-NA
         myspread$display[spreadrowcount]<-'break'
         spreadrowcount<-spreadrowcount+1 }
    myspread$skip_ref[spreadrowcount]<-myblock #not sure what this variable does! It is here for historic reasons
    myspread$ES[spreadrowcount]<-E
    if(E>0)
    {myspread$ANSWER[spreadrowcount] <-'Blue>Pink'}
    myspread$trial[spreadrowcount]<-thistrial
    
  #for each effect size, create dataset in mydf, with half Exp and half Control

mydf$Score <- rnorm(N,mean=E,sd=1) #take random normal sample with mean of E for experimentals
mydf$Score[1:(N/2)] <- rnorm((N/2),mean=0,sd=1) #take random normal sample with mean of zero for controls

# create data set which includes accumulating data for each level
startcol<- which(colnames(myspread)=='meanC1')
startmeans<- seq(startcol,startcol-1+(allN*4),4) #starting col for writing means for each N_size
EScol<-which(colnames(myspread)=='ES1') #start of ES column range
LLcol<-which(colnames(myspread)=='LL1') #although this could be computed, this is easier, and more explicit
tcol<-which(colnames(myspread)=='t1') #although this could be computed, this is easier, and more explicit
BFcol<-which(colnames(myspread)=='BF1') #although this could be computed, this is easier, and more explicit


  for (j in 1:allN){ 
   
    myN <- N_size[j]/2 #myN is sample size per group and varies from 10 to 320
       range1a<-1
      range1b<-myN #control range from 1:myN
      range2a<-(N/2)+1
      range2b<-N/2+myN #exp range 
# add means and sds for con then exp to spreadsheet
     myspread[spreadrowcount,startmeans[j]]<-mean(mydf$Score[range1a:range1b])
    myspread[spreadrowcount,(startmeans[j]+1)]<-sd(mydf$Score[range1a:range1b])
     myspread[spreadrowcount,(startmeans[j]+2)]<-mean(mydf$Score[range2a:range2b])
    myspread[spreadrowcount,(startmeans[j]+3)]<-sd(mydf$Score[range2a:range2b])
# add Effect size for this sample size to spreadsheet
      ESObs<-(-1)*effsize::cohen.d(mydf$Score[range1a:range1b],mydf$Score[range2a:range2b])$estimate
        myspread[spreadrowcount,(EScol+j-1)]<-ESObs
      # add Loglikelihood for this sample size to spreadsheet

       myse<-sqrt(sd(mydf$Score[range1a:range1b])/myN+
        sd(mydf$Score[range2a:range2b])/myN)
       meandiff<-mean(mydf$Score[range2a:range2b])-mean(mydf$Score[range1a:range1b])
       myspread[spreadrowcount,(EScol+j-1+allN)]<-log(dnorm(meandiff,mean=.3,sd=myse)/
                        dnorm(meandiff,mean=0,sd=myse))
       
     # add t-value for this sample size to spreadsheet
       myt<-t.test(x=mydf$Score[range1a:range1b],y = mydf$Score[range2a:range2b], paired=FALSE)
       myspread[spreadrowcount,(EScol+j-1+(2*allN))]<-myt$statistic
 
     # add BF for this sample size to spreadsheet
          bf = ttestBF(x = mydf$Score[range1a:range1b],y=mydf$Score[range2a:range2b], paired=FALSE)
         myspread[spreadrowcount,(EScol+j-1+(3*allN))] <-as.data.frame(bf)$bf
 
 #Add this bit of data to big_data, with sample size code (cnd_sample)
 #I.e. big_data has all samples on top of each other, so there will be repetitions as the data cumulates
 add_bit<-mydf[c(range1a:range1b,range2a:range2b),]
 add_bit$cnd_sample<-myN
 if (j==1){big_data<-add_bit}
   if(j>1){
   big_data<-rbind(big_data,add_bit)
   }
 
  } #all sample sizes complete

big_data$cnd_sample <- as.factor(big_data$cnd_sample)

filename<-paste0("E", E*10,'_',num2alpha(thistrial))
myspread$image1[spreadrowcount]<-paste0(filename,'.gif')
myspread$image1f[spreadrowcount]<-paste0(filename,'.png')
 makegif(big_data,E,filename) #take the current generated df, create gifs for samples, with final plot showin g effect size E.
}
myspread$display[nrow(myspread)]<-'end'
myspread$ANSWER[nrow(myspread)]<-NA

write.csv(myspread, "gifs_pngs/Debug.csv")
```

