---
title: "bayes_vs_effsize"
author: "DVM Bishop"
date: "30/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(psycho) #for dprime
```

## Script to explore ways of analysing data from probability game
Background and locations:
Game is here:  https://gorilla.sc/admin/project/7839#
Draft preregistration document is here:  https://osf.io/rnkhq/registrations?tab=drafts
This was also copied onto a word document, saved on Dropbox/oscci_ms_2016-/ongoing/jackie_game 



For each sample size of each trial, we have effect size and log likelihood.
Question will be: are people using these to make their judgements?
Does this change over course of the session?
Could it be influenced by reward schedule?

Simulated data were created by file 'make_gifs_datasim.R'.
We read in a large file created this way: bigBFdata.csv

```{r readtrialdata}
mydir<-'~/Dropbox/jackieT_project/bayesian attempt' 
mydir<-getwd()
mytrials<-read.csv(paste0(mydir,'/bigBFdata.csv'))
#check observed effect sizes - these confirm they are around .3, .5, and .8, and zero

mycheck<-aggregate(mytrials$obsES, list(mytrials$trueES), mean)
names(mycheck)<-c('True ES','Observed ES')
mycheck

mymeanC<-aggregate(mytrials$meanC, list(mytrials$trueES), mean)

mysdC<-aggregate(mytrials$meanC, list(mytrials$trueES), sd)
mymeanC$sd <- mysdC[,2]
names(mymeanC)<-c('True ES','meanC','semC')
mymeanC

mymeanE<-aggregate(mytrials$meanE, list(mytrials$trueES), mean)
names(mymeanE)<-c('True ES','meanE')
mysdE<-aggregate(mytrials$meanE, list(mytrials$trueES), sd)
mymeanE$sd <- mysdE[,2]
names(mymeanE)<-c('True ES','meanE','semE')
mymeanE
```
We can now see how accurate a subject would be if they adopted a particular strategy for responding.

I started looking at strategy based on effect size - i.e. compare means of C and E distributions, but if you know C has no effect, could also just ignore C and base strategy solely on mean of E.

We can use the mean for E and compute log likelihood that it came from E or C distribution.

```{r loglikelihood}
grain=.025 #specifies width of interval for computing Nexp and Nnull

mytrials$seE <- mytrials$sdE/sqrt(mytrials$Nsize)

mytrials$Nnull <- Nsample*(pnorm((mytrials$meanE+(grain/2)),0,mytrials$seE,lower.tail=T)-
                             pnorm((mytrials$meanE-(grain/2)),0,mytrials$seE,lower.tail=T))
    #use range of observed effect for each value of myint +/- .05
mytrials$Nexp <- Nsample*(pnorm((mytrials$meanE+(grain/2)),mytrials$trueES,mytrials$seE,lower.tail=T)-
                             pnorm((mytrials$meanE-(grain/2)),mytrials$trueES,mytrials$seE,lower.tail=T))
mytrials$LL <- log(mytrials$Nexp/mytrials$Nnull)

plot(mytrials$meanE,mytrials$LL,col=as.factor(mytrials$Nsize))
```

Consider the following strategy:
a. Response NO if the Effect Size is less than 0
b. Response YES if the Effect size is greater than .2
Otherwise wait to accumulate more evidence.
We can try this on our total file of data to see how accurate someone would be if they do this.

We need to add columns denoting Response (No=0, Yes=1, NA for Wait or if response already made on this trial)

We can then assign a score of correct/incorrect.
We need also to have a response rule for the case when all sample sizes have been seen and the criterion has not been met. This would be YES if effect size > .2 and NO otherwise.

```{r tryEScriterion}
mytrials$respES2<-NA #column to hold responses if effect size criterion is used
mytrials$corrES2<-NA#column to hold accuracy of response if effect size criterion is used
critES<-.2 #if obs effect is greater than this value, respond YES
critESn<-.2 #if obs effect is less than this value, respond NO
Nsizes<-6 #number of sample sizes that are presented on each trial

thisrow<-0
for (i in 1:(nrow(mytrials)/Nsizes)){
  critflag<-0 #this flag is set when conditions are met for a response; reset for each trial
  for (thissize in 1:Nsizes){
    thisrow<-thisrow+1
    if(critflag<1){ #no response is given yet, so proceed to check if criterion met
  if(mytrials$obsES[thisrow]>critES){ #does the observed effect size exceed critical effect size?
    critflag<-1
    mytrials$respES2[thisrow]<-1 #for this response, record YES, and set critflag to 1 to denote response made
  }
      if(mytrials$obsES[thisrow]<critESn){ #is observed effect size less than critESn?
    critflag<-1
    mytrials$respES2[thisrow]<-0 #for this response, record NO
      }
    } #proceed to this point if critflag is still zero
    if((thissize==Nsizes)&(critflag==0)){ #got to last size and still not met criterion; adopt default criterion; assume NO unless obs effect size is bigger than criterion effect size
         mytrials$respES2[thisrow]<-1 #for this response, record YES
         if(mytrials$trueES[thisrow]<critES){mytrials$respES2[thisrow]<-0}
    }
    #now decide if the response was correct or not
   if(!is.na(mytrials$respES2[thisrow])){
    trueeff<-0
    if(mytrials$trueES[thisrow]>0){trueeff<-1} #true effect size in population is greater than zero
    mytrials$corrES2[thisrow]<-0 #default response code is error
    if(mytrials$respES2[thisrow]==trueeff) {mytrials$corrES2[thisrow]<-1} #switch to correct if resp matches trueeff
   }
    }
}

myacc <-aggregate(mytrials$corrES2, by=list(mytrials$trueES),
  FUN=mean, na.rm=TRUE)
names(myacc)<-c('True effect size','p.Correct')

```

Seems an asymmetry with more false positives (really null but says yes) than misses (says null when real effect)

How does this compare with use of Bayes Factor.
To give some flexibility in selection of a criterion, will make prior chunk into a function

```{r flexcriterion}
myflex<-function(mytrials,respcol,corrcol,critcol,critvalue,critvalueN,Nsizes,sizecol){
  
  thisrow<-0
  for (i in 1:(nrow(mytrials)/Nsizes)){
    critflag<-0 #this flag is set when conditions are met for a response; reset for each trial
    for (thissize in 1:Nsizes){
      thisrow<-thisrow+1
      if(critflag==0){ #no response is given yet, so proceed to check if criterion met
        if(mytrials[thisrow,criterioncol]>critvalue){ #does the observed effect size exceed critical effect size?
          critflag<-1
          mytrials[thisrow,responsecol]<-1 #for this response, record YES, and set critflag to 1 to denote response made
        }
        if(mytrials[thisrow,criterioncol]<critvalueN){ #is observed effect size less than critESn?
          critflag<-1
          mytrials[thisrow,responsecol]<-0 #for this response, record NO
        }
      } #proceed to this point if critflag is still zero
      if((thissize==Nsizes)&(critflag==0)){ #got to last size and still not met criterion; adopt default criterion; assume NO unless obs effect size is bigger than criterion effect size
        mytrials[thisrow,responsecol]<-1 #for this response, record YES
        if(mytrials[thisrow,criterioncol]<critvalue){mytrials[thisrow,responsecol]<-0}
      }
      #now decide if the response was correct or not
      if(!is.na(mytrials[thisrow,responsecol])){ #ignore those where no response
        trueeff<-0 #check if true effect size is zero or greater
        if(mytrials$trueES[thisrow]>0){trueeff<-1} #true effect size in population is greater than zero
        mytrials[thisrow,correctcol]<-0 #default code is error
        if(mytrials[thisrow,responsecol]==trueeff) {mytrials[thisrow,correctcol]<-1} #switch to correct if resp matches trueeff
        mytrials[thisrow,sizecol]<-mytrials$Nsize[thisrow]
      }
    }
  }
  
  myacc <-aggregate(mytrials[,c(sizecol,correctcol)], by=list(mytrials$trueES),
  FUN=mean, na.rm=TRUE)
  names(myacc)<-c('True effect size','N','p.Correct')
  print(myacc)
  return(mytrials)
}
```

```{r dologBFcrit}
summaryperf<-data.frame(matrix(ncol = 7, nrow = 100))
colnames(summaryperf)<-c('type','critvalY','critvalN','p.corrYES','p.corrNO','meanNYES','meanNNO')
Nsizes<-6 #number of sample sizes that are presented on each trial
mytype<-'logBF'
critvals<-c(1,1.5,2,2.5,3,3.5,4) #all the critical values to be tested
critvalNs<-c(-3,-2.5,-2,-1.5,-1)#all the critical values for neg result to be tested
#comment out next 2 lines if you want to test all values - below are single values to test the function
critvals<-1
critvalNs<-(-1)
myrow<-which(is.na(summaryperf$type))[1]-1 #start writing data to first blank row
for(critvalue in critvals){
  for(critvalueN in critvalNs){

  mytrials$resplBF<-NA #column to hold responses if effect size criterion is used
mytrials$corrlBF<-NA#column to hold accuracy of response if effect size criterion is used
mytrials$size<-NA #col to hold size
responsecol<-which(names(mytrials)=='resplBF') #find column number for resp
correctcol<-which(names(mytrials)=='corrlBF') #find column number for correct
criterioncol<-which(names(mytrials)=='logBFH1') #column that holds the key statistic
sizecol<-which(names(mytrials)=='size') #holds N at point when response made
 myrow<-myrow+1
 summaryperf$type[myrow]<-mytype
 summaryperf$critvalY[myrow]<-critvalue
  summaryperf$critvalN[myrow]<-critvalueN
 mytrials<-myflex(mytrials,responsecol,correctcol,criterioncol,critvalue,critvalueN,Nsizes,sizecol)
 w<-intersect(which(mytrials$trueES>0),which(!is.na(mytrials$corrlBF)))
 summaryperf$p.corrYES[myrow]<-sum(mytrials[w,correctcol],na.rm=TRUE)/(nrow(mytrials)/Nsizes/2)
  summaryperf$p.corrNO[myrow]<-sum(mytrials[-w,correctcol],na.rm=TRUE)/(nrow(mytrials)/Nsizes/2)
 summaryperf$meanNYES[myrow]<-mean(mytrials$Nsize[w],na.rm=TRUE)
 summaryperf$meanNNO[myrow]<-mean(mytrials$Nsize[-w],na.rm=TRUE)
}
}

mytype<-'ES'

myrow<-which(is.na(summaryperf$type))[1]-1 #start writing data to first blank row
critvals<-c(.1,.15,.2,.25,.3,.35,.4)
critvalNs<-c(-.3,-.25,-.2,-.15,-.1,-.05,0)
#comment out next 2 lines for full run: lines below used for testing
critvals<-.3
critvalNs<-(-.1)
for(critvalue in critvals){ #if obs effect is greater than this value, respond YES
for(critvalueN in critvalNs){ #if obs effect is less than this value, respond NO

  mytrials$respES<-NA #column to hold responses if effect size criterion is used
mytrials$corrES<-NA#column to hold accuracy of response if effect size criterion is used
mytrials$size<-NA#col to hold N at point of response
responsecol<-which(names(mytrials)=='respES') #find column number for resp
correctcol<-which(names(mytrials)=='corrES') #find column number for correct
sizecol<-which(names(mytrials)=='size')
criterioncol<-which(names(mytrials)=='obsES') #column that holds the key statistic

 myrow<-myrow+1
 summaryperf$type[myrow]<-mytype
 summaryperf$critvalY[myrow]<-critvalue
  summaryperf$critvalN[myrow]<-critvalueN
 mytrials<-myflex(mytrials,responsecol,correctcol,criterioncol,critvalue,critvalueN,Nsizes,sizecol)
 w<-intersect(which(mytrials$trueES>0),which(!is.na(mytrials[,correctcol])))
 summaryperf$p.corrYES[myrow]<-sum(mytrials[w,correctcol],na.rm=TRUE)/(nrow(mytrials)/Nsizes/2)
  summaryperf$p.corrNO[myrow]<-sum(mytrials[-w,correctcol],na.rm=TRUE)/(nrow(mytrials)/Nsizes/2)
 summaryperf$meanNYES[myrow]<-mean(mytrials$size,na.rm=TRUE)
 summaryperf$meanNNO[myrow]<-mean(mytrials$size,na.rm=TRUE)
}
}

 
```


I will simulate choices by a fictitious subject: these will give (a) the N at which the response was made, and (b) the choice (1 or 0).

From this, we can get an obs effect size and Bayes Factor corresponding to that choice.

In this first attempt, the simulation is just random: the subject stops at a random point and selects true or false at random

```{r fictitiousdata}
Elist<-c(3,5,8,0) #effect size x 10
nsets<-c(6,6,6,18) #N sets at each effect size
N_size <- c(20, 40, 80, 160, 320, 640) #sample sizes used
earning<-c(-10,10,-1) #reward schedule: loss if wrong, gain if right, loss per N level

mydf<-data.frame(matrix(nrow=(sum(nsets)),ncol=12))
names(mydf)<-c('run','lettercode','trueES','sel_N','response','obsES','t','logBFH1','correct','earn','penalty','balance')

newdf <- mytrials[c(rep(FALSE,5),TRUE), ] #just using existing columns in mytrials to populate start of mydf
mydf[,1:3]<-newdf[,c(1,2,4)]
mydf$sel_N<-sample(N_size,nrow(mydf),replace=TRUE) #random selection of a sample size
mydf$response<-sample(c(0,1),nrow(mydf),replace=TRUE) #random selection of a response
#we now need to look up the obs effect size and BF for this response pattern

for (i in 1:nrow(mydf)){
  thisrun<-mydf$run[i]
  thisletter<-mydf$lettercode[i]
  thisN<-mydf$sel_N[i]
w<-intersect(intersect(which(mytrials$run==thisrun),which(mytrials$Nsize==thisN)),which(mytrials$lettercode==thisletter)) #Phew! I tried filter and could not get it to work!
  mydf$obsES[i]<-mytrials$obsES[w]
  mydf$logBFH1[i]<-mytrials$logBFH1[w]
  mydf$t[i]<-mytrials$t[w]
}
```
Do computation of cost/benefit in a separate function

```{r costs}
mycost<-function(mydf){
for (i in 1:nrow(mydf)){
  mydf$correct[i]<-0 #default is error
  if ((mydf$trueES[i]>0)&(mydf$response[i]==1))
      {mydf$correct[i]<-1}
  if ((mydf$trueES[i]==0)&(mydf$response[i]==0))
    {mydf$correct[i]<-1}  
  mydf$earn_cost[i]<- earning[1]
  if (mydf$correct[i]==1)
  {mydf$earn_cost[i]<- earning[2]}
  w<-which(N_size==mydf$sel_N[i])
  mydf$penalty[i]<-(w-1)*earning[3] #can set earning [3] to zero for no cost of waiting
}
  return(mydf)
  }
```

```{r countcosts}
mydf<-mycost(mydf)
```
We can now check whether selection of a response is related to either effect size or t-value or Bayes Factor


```{r compareresponse}

t.test(abs(mydf$logBFH1)~mydf$response) #use absolute value of bayes factor
t.test(mydf$obsES~mydf$response)
t.test(mydf$t~mydf$response)



```

# Finding optimal strategy

Possible rule. 
Decide if log Bayes is greater than absolute value of a critical level.
If it doesn't reach that level by biggest N, then go by whether effect size is > .2

Alternative rule: just go by whether effect size is > critical ES

Or whether t-value is > 1.96

We'll start by making a dataframe to save results for different settings


Now we'll create a function that for trying different settings for criterion to make decision, and save the results in mydfsummary

```{r trycriterion}
mytry<-function(mytrials,mydf,thiscol,critvalue,thiscorrectcol,mydefaultcut){
  #we need data from mytrials, plus the original dataframe (mydf) as a dummy to overwrite
  #mycol specifies which column is used to decide on response
  #mycutoff specifies which cutoff
  #mycorrectcol is the default column for decision if cutoff not reached - typically ES
  #mydefaultcut is cutoff used with default column; typically .2

mycol<-which(names(mytrials)==thiscol) #find the column with the name from thiscol
mycorrectcol<-which(names(mytrials)==thiscorrectcol) #find the column with the name from thiscol

mydf1<-mydf #start by copying mydf - we will overwrite the selections etc
mydf1[,4:11]<-NA 
mydf1$confident<-0 #this is a new column used to record whether the cutoff led to a decision.  If not, then have to rely on default.
ntrial<-max(mytrials$run)

for (myt in 1:ntrial){
  startrow<-1+(myt-1)*6
  for (thisrow in (startrow+5):startrow){ #go backwards to find earliest N meeting criteria
    myfound<-0 #marker for whether a value meeting criterion is found
    if(mytrials[thisrow,mycol]>critvalue){ #e.g. check if abs(logBF) greater than 3, 
      myfound<-1
      mydf1$confident[myt]<-1 #we have a value meeting criterion
    mydf1$response[myt]<-1 #respond TRUE
    mydf1$sel_N[myt]<-mytrials$Nsize[thisrow] #record the N for this selection
    mydf1$obsES[myt]<-mytrials$obsES[thisrow] 
    mydf1$logBFH1[myt]<-mytrials$logBFH1[thisrow]
    mydf1$t[myt]<-mytrials$t[thisrow]
    }
     if(mytrials[thisrow,mycol]< -(critvalue)){ #mirror image for null hypoth
      myfound<-1
       mydf1$confident[myt]<-1
       mydf1$response[myt]<-0
    mydf1$sel_N[myt]<-mytrials$Nsize[thisrow]
    mydf1$obsES[myt]<-mytrials$obsES[thisrow]
    mydf1$logBFH1[myt]<-mytrials$logBFH1[thisrow]
    mydf1$t[myt]<-mytrials$t[thisrow]}

  } #finish checking all rows

      
  if(mydf1$confident[myt]==0){ 
    thisrow<-startrow+5 #did not find any case of meeting criterion, so select highest N (640)
    #for these trials, we we use mycorrectcol and mydefaultcut to determine response
    mydf1$response[myt]<-0 #respond FALSE
    mydf1$sel_N[myt]<-mytrials$Nsize[thisrow]
    mydf1$obsES[myt]<-mytrials$obsES[thisrow]
      mydf1$t[myt]<-mytrials$t[thisrow]
    mydf1$logBFH1[myt]<-mytrials$logBFH1[thisrow]
    if(mytrials[thisrow,mycorrectcol]>mydefaultcut){
      mydf1$response[myt]<-1 #respond TRUE
    }
  }
}
mydf1$correct<-1
w<-c(which((mydf1$response==1)&(mydf1$trueES==0)),which((mydf1$response==0)&(mydf1$trueES > 0)))
if(length(w)>0){
mydf1$correct[w]<-0
}
return(mydf1)
}


```

```{r populatesummary}
#Use output from mytry function to summarise results and record in mydfsummary
mypop<-function(thisdf,mydfsummary,sumrow){

myt1<-t.test(thisdf$logBFH1~thisdf$response)$statistic #use absolute value of bayes factor
myt2<-t.test(thisdf$obsES~thisdf$response)$statistic
myt3<-t.test(thisdf$t~thisdf$response)$statistic
mydfsummary$type[sumrow]<-thiscol
mydfsummary$cutoff[sumrow]<-critvalue
mydfsummary$p.confident[sumrow]<-mean(thisdf$confident)
mydfsummary$default[sumrow]<-thiscorrectcol
mydfsummary$default.setting[sumrow]<-mydefaultcut
mydfsummary$p.corr[sumrow]<-mean(thisdf$correct)
mydfsummary$avgN[sumrow]<-mean(thisdf$sel_N)
mydfsummary$t.ES[sumrow]<-myt2
mydfsummary$t.t[sumrow]<-myt3
mydfsummary$t.BF[sumrow]<-myt1
mydfsummary$totearn[sumrow]<-sum(earning[1]*(1-thisdf$correct))+sum(earning[2]*thisdf$correct)
mydfsummary$totalpenalty[sumrow]<-sum(thisdf$penalty)
mydfsummary$balance[sumrow]<-mydfsummary$totearn[sumrow]+mydfsummary$totalpenalty[sumrow]
return(mydfsummary)

}
```

```{r makesummarydf}
mydfsummary<-data.frame(matrix(nrow=100,ncol=13))
names(mydfsummary)<-c('type','cutoff','p.confident','default','default.setting','p.corr','avgN','t.ES','t.t','t.BF','totearn','totalpenalty','balance')
#As well as saving % correct, we'll save the t-value for comparing ES, t-value and BF for 'yes' and 'no' responses. These should mirror which variable used for comparison
```

```{r trycutoffs}

bflist<-c(1,1.5,2,2.5,3)
thiscol<-'logBFH1'
thiscorrectcol<-'obsES' #this is the column used to make decision if mycol criterion not met
mydefaultcut<-.2 #cutoff for mycorrectcol
for (critvalue in bflist){
thisdf<-mytry(mytrials,mydf,thiscol,critvalue,thiscorrectcol,mydefaultcut)
thisdf<-mycost(thisdf)
 sumrow<-min(which(is.na(mydfsummary$type))) #first unpopulated row
mydfsummary<-mypop(thisdf,mydfsummary,sumrow)
}

eslist<-c(.2,.3,.5,.8)
thiscol<-'obsES'
for (critvalue in eslist){
thisdf<-mytry(mytrials,mydf,thiscol,critvalue,thiscorrectcol,mydefaultcut)
thisdf<-mycost(thisdf)
 sumrow<-min(which(is.na(mydfsummary$type))) #first unpopulated row
mydfsummary<-mypop(thisdf,mydfsummary,sumrow)

}

tlist<-c(1.65,1.96,2.06,3)
thiscol<-'t'
for (critvalue in tlist){
thisdf<-mytry(mytrials,mydf,thiscol,critvalue,thiscorrectcol,mydefaultcut)
thisdf<-mycost(thisdf)
 sumrow<-min(which(is.na(mydfsummary$type))) #first unpopulated row
mydfsummary<-mypop(thisdf,mydfsummary,sumrow)
}
```



